Predicting instabilities of a tuneable ring laser with an iterative map model

### 2020-11-29 Papers/Journal/Tuneable_Lasers.tex
Fixed figure justification

\subsection{Combining the Effects}
\label{sec:effects}

@@

We now turn our attention to a more quantitative analysis, and characterize the stability of a laser by exploring the parameter space. We focus on the $s$--$b$ plane in particular, since $a$ and $h$ effect only the amplitude of the pulse, whereas the interplay of dispersion and nonlinearity gives rise to the instabilities. Additionally, as mentioned, a conservative initial energy will effect the dynamics much less than a more intense pulse since the nonlinearity will be negligible while initially coming to equilibrium. We examine the relative change in the pulse's envelope between consecutive round trips of the cavity to identify the stability of a pulse. We compute the error as
\begin{align}
	\Delta = \frac{\| |\mathcal{L}^n(A_0)| - |\mathcal{L}^{n-1}(A_0)| \|_2}{\| \mathcal{L}^{n-1}(A_0) \|_2},
	\label{eq:error}
\end{align}
where
\begin{align}
	\| f \|_2^2 \coloneqq \int_{-\infty}^\infty |f|^2 \, \df T,
\end{align}
and we choose $n$ to be sufficiently large to guarantee the pulse reaches equilibrium or deteriorates because of modulation instability---we use $n = 500$ in our experiments. We take the modulus of the pulse since we are interested in the evolution of only the envelope because of the auto-correlation methods used for experimental measurements. We plot the error in Fig.~\ref{fig:error}. Two distinct regions split the parameter space, divided by an incredibly sharp, and complicated boundary. In the upper-left region, the error is $\bigO{1}$; here, the nonlinearity is too strong, and the SPM induces modulation instability. In the lower-right region we find the opposite behaviour---the pulse reaches an equilibrium state and is stable. The increase in dispersion, coupled with modulation, allows the laser to balance the nonlinear effects. The error in this region is on the order of machine precision, suggesting the pulse has come to equilibrium.

@@

Finally, we briefly investigate the effect the order of the components has on the dynamics. In Section~\ref{sec:model}\ref{sec:effects} we justified the order of the gain, nonlinearity, and loss, but, just assumed dispersion preceded modulation. Here, we consider the effect of swapping the dispersion and modulation components. The overall structure of the energy landscape (Fig.~\ref{fig:energyswitch}) is the same as in Fig.~\ref{fig:energy}. In the previous case, with modulation following dispersion, more energy is removed from the pulse tails. Whereas, having modulation first, more energy remains in the cavity. The additional energy makes the laser more powerful than in the previous case, which in turn increases the size of the unstable region. Comparing with Fig.~\ref{fig:energy}, the unstable upper-left region spreads farther rightward, and the energy is larger in the stable region---especially away from the origin. Once again, the energy at the origin is 2.5487, as we are effectively removing the CFBG in the limit $s \rightarrow 0$.

>>>>

\subsection{Combining the Component Maps}
\label{sec:effects}

@@

We now turn our attention to a more quantitative analysis, and characterize the stability of a laser by exploring the parameter space. We focus on the $s$--$b$ plane in particular, since $a$ and $h$ affect only the amplitude of the pulse, whereas the interplay of dispersion and nonlinearity gives rise to the instabilities. Additionally, as mentioned, a conservative initial energy will affect the dynamics much less than a more intense pulse since the nonlinearity will be negligible while initially coming to equilibrium. We examine the relative change in the pulse's envelope between consecutive round trips of the cavity to identify the stability of a pulse. We compute the error as
\begin{align}
	\Delta = \frac{\| |\mathcal{L}^n(A_0)| - |\mathcal{L}^{n-1}(A_0)| \|_2}{\| \mathcal{L}^{n-1}(A_0) \|_2},
	\label{eq:error}
\end{align}
where
\begin{align}
	\| f \|_2^2 \coloneqq \int_{-\infty}^\infty |f|^2 \, \df T,
\end{align}
and we choose $n$ to be sufficiently large to guarantee the pulse reaches equilibrium or deteriorates because of modulation instability---we use $n = 500$ in our experiments. We take the modulus of the pulse since we are interested in the evolution of only the envelope because of the auto-correlation methods used for experimental measurements. We plot the error in Fig.~\ref{fig:error}. Two distinct regions split the parameter space, divided by an incredibly sharp, and complicated boundary. In the upper-left region, the error is $\bigO{1}$; here, the nonlinearity is too strong, and the SPM induces modulation instability. In the lower-right region we find the opposite behaviour---the pulse reaches an equilibrium state and is stable. The increase in dispersion, coupled with modulation, allows the laser to balance the nonlinear effects. The error in this region is on the order of machine precision, suggesting the pulse has come to equilibrium.

@@

Finally, we briefly investigate the affect the order of the components has on the dynamics. In Section~\ref{sec:model}\ref{sec:effects} we justified the order of the gain, nonlinearity, and loss, but, just assumed dispersion preceded modulation. Here, we consider the effect of swapping the dispersion and modulation components. The overall structure of the energy landscape (Fig.~\ref{fig:energyswitch}) is the same as in Fig.~\ref{fig:energy}. In the previous case, with modulation following dispersion, more energy is removed from the pulse tails. Whereas, having modulation first, more energy remains in the cavity. The additional energy makes the laser more powerful than in the previous case, which in turn increases the size of the unstable region. Comparing with Fig.~\ref{fig:energy}, the unstable upper-left region spreads farther rightward, and the energy is larger in the stable region---especially away from the origin. Once again, the energy at the origin is 2.5487, as we are effectively removing the CFBG in the limit $s \rightarrow 0$.
